# Practical 4
## Reference classes

```{r,P2, echo=FALSE}
## Solutions ##
randu = setRefClass("randu", 
                    fields = list(calls = "numeric", 
                                  seed = "numeric", 
                                  state = "numeric"))
randu$methods(get_state = function() state)
randu$methods(set_state = function(initial) state <<- initial)
randu$methods(get_seed = function() seed)
randu$methods(get_num_calls = function() calls)
randu$methods(r = function() {
  calls <<- calls + 1
  state <<- (65539 * state) %% 2^31
  return(state / 2^31)
})
```

The example in the notes created a random number generator using a reference class.

- Reproduce the `randu` generator from the notes and make sure that it works as advertised.^[The reference class version, not the function closure generator.]
- When we initialise the random number generator, the very first state is called the `seed`. Store this variable and create a new function called `get_seed` that will return the initial seed, i.e.^[Reference classes also have an initialise method - that way we would only specify the seed and would then initialise the other variables. I'll give you an example in the solutions.]
```{r,}
r = randu(calls = 0, seed = 10, state = 10)
r$r()
r$get_state()
r$get_seed()
```
```{r}
##Solutions - see below
```
- Create a variable that stores the number of times the generator has been called. You should be able to access this variable with the function `get_num_calls`
```{r,}
r = randu(calls = 0, seed = 10, state = 10)
r$get_num_calls()
r$r()
r$r()
r$get_num_calls()
```


```{r,P2}
```

## Solutions

Solutions are contained within the course package
```{r,eval=FALSE, echo = TRUE}
library("jrAdvanced")
vignette("solutions4", package = "jrAdvanced")
```