## S3 objects

1. Following the cohort example in the notes, suppose we want to create a `mean` method.
2. List all S3 methods associated with the`mean` function.
```{r }
methods("mean")
```
3. Examine the source code of `mean`. 
```{r results="hide"}
body("mean")
```
4. What are the arguments of `mean`?
```{r }
args("mean")
```
5. Create a function called `mean.cohort` that returns a vector containing the mean weight and mean height. Ensure that you can pass in the standard `mean` arguments, i.e.`na.rm`.`
```{r results='hide'}
mean.cohort = function(x, ...) {
  m1 = mean(x$details[, 1], ...)
  m2 = mean(x$details[, 2], ...)
  return(c(m1, m2))
}
```
6. Let's now make a similar function for the standard deviation:
    - Look at the arguments of the `sd` function. 
    - Create a default `sd` function. Look at `cor.default` in the notes for a hint.
    - Create an function call `sd.cohort` that returns a vector containing the weight and height standard deviation. Ensure that you can pass in the standard `sd` arguments, i.e. `na.rm`.`
```{r}
args(sd)
sd = function(x, ...) UseMethod("sd")
sd.default = function(x, ...) stats::sd(x, ...)
sd.cohort = function(x, ...) {
  s1 = sd(x$details[, 1], ...)
  s2 = sd(x$details[, 2], ...)
  return(c(s1, s2))
} 
```
7. We want to create a `summary` method for the `cohort` class. When the `summary` function is called on a `cohort` object it should call the base `summary` on the `details`  element:
    - Use the `body` function to check if the function is already a generic function.
    - Use the `args` function to determine the arguments.
    - Create a `summary.cohort` function
```{r }
## summary is already a generic
body(summary)

## Match the args
args(summary)

## Function
summary.cohort = function(object, ...) summary(object$details, ...)
```
8. Create a `hist` method for the `cohort` class. When the `hist` function is called on a `cohort` object, it should produce a single plot showing two histograms - one for height and another for weight.
```{r }
## hist is already a generic
body(hist)

## Match the args
args(hist)

## Function
hist.cohort = function(x, ...) {
  op = par(mfrow = c(1, 2))
  hist(x$details[, 1], main = "Weight")
  hist(x$details[, 2], main = "Height")
  par(op)
}
```
12. Create a `[` method for the `cohort` class. This method should return a `cohort` object, but with the relevant rows sub setted. For example, if `cc` was a cohort object, then
```{r, eval=FALSE, echo = TRUE}
cc[1:3, ]
```
\noindent would return the first three rows of the data frame.
```{r }
## Lots of methods available. 
methods("[")

## Examine [.data.frame
args("[.data.frame")

"[.cohort" = function(x, ...) {
  x$details = x$details[...]
  x
}
```
13. Create a `[<-` method for the `cohort` class. This method should allow us to replace values in the `details` data frame, i.e.
```{r, eval=FALSE, echo = TRUE}
cc[1, 1] = 10
```
```{r , eval = FALSE}
## Lots of methods available. 
methods("[<-")

## Examine [.data.frame
args("[<-.data.frame")

"[<-.cohort" = function(x, i, j, value) {
  x$details[i, j] = value
  x
}
cc[1:3, ] = 55
```


## Solutions

Solutions are contained within the course package
```{r, eval=FALSE, echo = TRUE}
library("jrAdvanced")
vignette("solutions2", package = "jrAdvanced")
```